---
title: "LMIS_Titles"
author: "Sam Neylon"
date: "August 4, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#### Using "here" to keep project directory as working directory
#### NOTE: Rmarkdown will try to make the file it is inside the working directory, but "here" should fix it.

library(here)
library(tidyverse)
library(data.table)

```
  
# Import

## fread

```{r eval=TRUE}

# Import/Rename columns

nyc_data <- fread(here("data/ALL_Citywide_Payroll_Data__Fiscal_Year_.csv"), col.names = c("FY", "payroll_num", "agency", "last", "first", "ME", "start_date", "borough", "title", "leave_status", "base_salary", "pay_basis", "reg_hours", "reg_grosspaid", "OT_hrs", "OT_paid", "total_otherPay"))

# Rename columns - Filter for FY2019-2020

nyc_data <- nyc_data %>% select(FY, agency, title, base_salary, pay_basis, reg_hours, reg_grosspaid, OT_hrs, OT_paid, total_otherPay) %>% 
  filter(FY > 2018)


```

# Clean

* Filter out zero reg_grosspaid and zero reg_hours
* "gross_wOT": column adding up Gross Paid and total overtime (OT) paid
* "hrly": Calculate "hourly" (not official hourly rate) by dividing Gross Paid by Regular (annual) hours

* Some job titles have asterisk or question marks - after cursory look at the data, I don't believe they are substantially different, so I delete these symbols.

```{r eval=TRUE}

nyc_data <- nyc_data %>%
  filter(reg_grosspaid > 0 & reg_hours > 0) %>% 
  mutate(
  gross_wOT = reg_grosspaid + OT_paid,
  hrly = reg_grosspaid / reg_hours
)

nyc_data <- nyc_data %>% mutate(
  title = gsub("\\*", "", title),
  title = gsub("\\?", "", title)
)

```

# Summary Table

I group rows by job title, so that each statistic is calculated on all the observations within a title. This pools multiple fiscal years, which I why I filtered for fiscal years I was interested in earlier.

```{r eval=TRUE}

nyc_group <- nyc_data %>% group_by(title)

nyc_wages <- nyc_group %>% summarise(
  tot_emp = n(),
  avg_pay = mean(reg_grosspaid),
  med_pay = median(reg_grosspaid),
  avg_pay = mean(reg_grosspaid),
  pay_q25 = quantile(reg_grosspaid, .25),
  pay_q75 = quantile(reg_grosspaid, .75),
  avg_hourly = mean(hrly),
  med_hourly = median(hrly),
  hrly_q25 = quantile(hrly, .25),
  hrly_q75 = quantile(hrly, .75),
  avg_hours = mean(reg_hours),
  med_hours = median(reg_hours),
  q25_hours = quantile(reg_hours, .25),
  q75_hours = quantile(reg_hours, .75),
  OT_avg_pay = mean(gross_wOT),
  OT_med_pay = median(gross_wOT),
  OT_avg_pay = mean(gross_wOT),
  OT_pay_q25 = quantile(gross_wOT, .25),
  OT_pay_q75 = quantile(gross_wOT, .75)
)

rm(nyc_group)

```
 
# Export

```{r eval=FALSE}

write_csv(nyc_wages, "nyc_wages_FY2019-20_08-05-21.csv")

```

# Data Exploration

I want to see what difference "per Hour" vs "per Annum" makes...

```{r eval=FALSE}

nyc_basis <- nyc_data %>% group_by(pay_basis)

nyc_basis <- nyc_basis %>% summarise(
  tot_emp = n(),
  avg_pay = mean(reg_grosspaid),
  med_pay = median(reg_grosspaid),
  avg_pay = mean(reg_grosspaid),
  pay_q25 = quantile(reg_grosspaid, .25),
  pay_q75 = quantile(reg_grosspaid, .75),
  avg_hourly = mean(hrly),
  med_hourly = median(hrly),
  hrly_q25 = quantile(hrly, .25),
  hrly_q75 = quantile(hrly, .75),
  OT_avg_pay = mean(gross_wOT),
  OT_med_pay = median(gross_wOT),
  OT_avg_pay = mean(gross_wOT),
  OT_pay_q25 = quantile(gross_wOT, .25),
  OT_pay_q75 = quantile(gross_wOT, .75)
)


```

# Outliers

```{r eval=FALSE}

hist(nyc_data$reg_grosspaid)

```




