---
title: "LMIS_Titles"
author: "Sam Neylon"
date: "August 4, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#### Using "here" to keep project directory as working directory
#### NOTE: Rmarkdown will try to make the file it is inside the working directory, but "here" should fix it.

library(here)
library(tidyverse)
library(data.table)

```

# Emp by Agency by Year

This file produces employee counts using the NYC OpenData payroll information.

NYC OpenData: https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e

We are comparing with headcount data from the Citizens Budget Commission as a robustness check of our data - https://cbcny.org/research/nyc-employee-headcount

*NOTE* This code is based on code which summarizes wages, and I left much of that code in.

Goal: Create a table which has employment in each agency for each fiscal year
  * Row: Agency
  * Column: Fiscal Year

# Import

## fread

```{r eval=TRUE}

# Import/Rename columns

nyc_data <- fread(here("data/ALL_Citywide_Payroll_Data__Fiscal_Year_.csv"), col.names = c("FY", "payroll_num", "agency", "last", "first", "ME", "start_date", "borough", "title", "leave_status", "base_salary", "pay_basis", "reg_hours", "reg_grosspaid", "OT_hrs", "OT_paid", "total_otherPay"))

# Rename columns - Filter for FY2020

nyc_data <- nyc_data %>% select(FY, agency, title, base_salary, pay_basis, reg_hours, reg_grosspaid)


```

# Clean

* Filter out zero reg_grosspaid and zero reg_hours
* "hrly": Calculate "hourly" (not official hourly rate) by dividing Gross Paid by Regular (annual) hours
* Create an adjusted 2021 wage by multiplying regular gross paid by 1.7%

* Some job titles have asterisk or question marks - after cursory look at the data, I don't believe they are substantially different, so I delete these symbols.

```{r eval=TRUE}

nyc_data <- nyc_data %>%
  filter(reg_grosspaid > 0 & reg_hours > 0) %>% 
  mutate(
  hrly = reg_grosspaid / reg_hours,
  reg_grosspaid21 = reg_grosspaid * 1.017
)

nyc_data <- nyc_data %>% mutate(
  title = gsub("\\*", "", title),
  title = gsub("\\?", "", title)
)

```

# Grouping Data

In order to 'summarise' I will first need to group by agency and year

```{r eval=TRUE}

nyc_group <- nyc_data %>% group_by(agency, FY)

nyc_emp <- nyc_group %>% summarise(
  tot_emp = n())

rm(nyc_group)

```

## Pivot

This created a table with a row for every agency-year. I will now 'pivot_wider' to put years into columns:

```{r eval=TRUE}

nyc_emp <- nyc_emp %>% pivot_wider(names_from = FY, values_from = tot_emp)

```

 
# Export

*NOTE* This code creates a csv. In the final versions I upload, I open the csv in Excel, convert to xlsx, format the data as a table, re-format text so currency variables and numbers are formatted correctly, move the file to NYC_Payroll/output, then upload to SharePoint.

```{r eval=FALSE}

write_csv(nyc_emp, "Emp_byAgency_byFY.csv")

```





