---
title: "Titles by Agency"
author: "Sam Neylon"
date: "August 4, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#### Using "here" to keep project directory as working directory
#### NOTE: Rmarkdown will try to make the file it is inside the working directory, but "here" should fix it.

library(here)
library(tidyverse)
library(data.table)

```

# FY20-21(adjusted) - Titles by Agency

This file produces a summary statistics for titles in the NYC OpenData payroll information.

NYC OpenData: https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e

This file summarizes pay data for job titles *within* agencies.

Notes:
  * This file uses the raw data for FY2020
  * This file then creates 2021 statistics by *1.7%*
    * This is in line with the adjustments made by the OEWS for state and local government workers: https://statistics.labor.ny.gov/lstechoes.shtm
  * In line with OEWS statistics, this file estimates two wage statistics:
    1. "Entry" - this is the average wage for the bottom 1/3 of wage earners in that title
    2. "Experienced" - this is the average wage for the top 2/3 of wage earners in that title.
  
# Import

## fread

```{r eval=TRUE}

# Import/Rename columns

nyc_data <- fread(here("data/ALL_Citywide_Payroll_Data__Fiscal_Year_.csv"), col.names = c("FY", "payroll_num", "agency", "last", "first", "ME", "start_date", "borough", "title", "leave_status", "base_salary", "pay_basis", "reg_hours", "reg_grosspaid", "OT_hrs", "OT_paid", "total_otherPay"))

# Rename columns - Filter for FY2020

nyc_data <- nyc_data %>% select(FY, agency, title, base_salary, pay_basis, reg_hours, reg_grosspaid) %>% 
  filter(FY == 2020)


```

# Clean

* Filter out zero reg_grosspaid and zero reg_hours
* "hrly": Calculate "hourly" (not official hourly rate) by dividing Gross Paid by Regular (annual) hours
* Create an adjusted 2021 wage by multiplying regular gross paid by 1.7%

* Some job titles have asterisk or question marks - after cursory look at the data, I don't believe they are substantially different, so I delete these symbols.

```{r eval=TRUE}

nyc_data <- nyc_data %>%
  filter(reg_grosspaid > 0 & reg_hours > 0) %>% 
  mutate(
  hrly = reg_grosspaid / reg_hours,
  reg_grosspaid21 = reg_grosspaid * 1.017,
  hrly21 = reg_grosspaid21 / reg_hours
)

nyc_data <- nyc_data %>% mutate(
  title = gsub("\\*", "", title),
  title = gsub("\\?", "", title)
)

```

# Annualizing Wages

There is one important issue with using "grosspaid" in this dataset. This is that if employees leave before a full year, their grosspaid will be very low, and thus will drag down the summary statistics that we are calculating.

We could simply take the grosspaid and divide it by the hours worked in a year (reg_grosspaid / reg_hours), which is how we calculate the "hrly" variable, and annualize it by multiplying by a set number of annual hours. 

*However*, some positions actually *are* seasonal, or work less than 2,080 hours a year (40 hrs a week * 52 weeks), so this would remove important information about the wages of certain titles.

We have settled on the following solutions:

  Seasonal: Through reading qualitative sources (union job descriptions, civil service listings) we may determine that a particular job title is seasonal. If so, we will treat it as seasonal in code which is tailored to a specific cluster of job titles (see "Parks_wages_titleClusters").

## Everything else: 

* Create new variable "FT_hours"
  * Calculate the 75th percentile of annual hours in a given Fiscal Year for a given Title.
  * If the 75th percentile is <= 1,820 hours, then set the FT_hours at 1,820 for this row (person-year).
  * If the 75th percentile is > 1,820 hours, then set FT_hours for this row to the 75th percentile for the title.
  
* Create new variable "FTE_pay"
  * If reg_hours >= FT_hours then set FTE_pay to reg_grosspaid
  * Else set FTE_pay to reg_grosspaid / reg_hours x FT_hours


```{r eval=TRUE}

nyc_data <- nyc_data %>% group_by(title) %>% mutate(hrq75 = quantile(reg_hours, .75)) %>% ungroup()

nyc_data <- nyc_data %>% mutate(FT_hours = ifelse(hrq75 <= 1820, 1820, hrq75),
                                FTE_pay = ifelse(reg_hours >= FT_hours, reg_grosspaid, 
                                                 hrly * FT_hours),
                                FTE_pay21 = FTE_pay * 1.017)

```

# Entry/Experienced

*NOTE* Using full-time equivalent pay (see above)

In line with OEWS statistics, this file estimates two wage statistics:
  1. "Entry" - this is the average wage for the bottom 1/3 of wage earners in that title
  2. "Experienced" - this is the average wage for the top 2/3 of wage earners in that title.
  
## First Tertile
  
  * First, calculate where the first third of wages within each title falls (the first tertile), and add this as a variable for each observation
  * Second, create two logical variables. Variable "Entry" is TRUE if a given worker's regular gross pay is less than or equal to the first tertile and FALSE if not. For "Exper" (experienced), TRUE if more than first tertile, FALSE if not.


```{r eval=TRUE}

nyc_group <- nyc_data %>% group_by(agency, title)

nyc_group <- nyc_group %>% mutate(tert = quantile(FTE_pay, 1/3), tert21 = quantile(FTE_pay21, 1/3)) %>% 
  mutate(EntryTERT = (FTE_pay <= tert),
         ExperTERT = (FTE_pay > tert),
         EntryTERT21 = (FTE_pay21 <= tert21),
         ExperTERT21 = (FTE_pay21 > tert21))

```

# Conditional Summaries

Now that I have a logical vector for Entry and Experienced, I can multiply this by the individual values (see code below) so my summaries are only for particular workers within a title.

NOTE: I have commented out two lines which gave me a count of how many Entry and Experienced in each position (I did this as a check on my work)

```{r eval=TRUE}

nyc_wages <- nyc_group %>% summarise(
  tot_emp = n(),
  #entry_emp = sum(EntryTERT),
  #exper_emp = sum(ExperTERT),
  avg_pay = mean(reg_grosspaid),
  med_pay = median(reg_grosspaid),
  entry = mean(FTE_pay[EntryTERT == TRUE]),
  exper = mean(FTE_pay[ExperTERT == TRUE]),
  avg_pay21 = mean(reg_grosspaid21),
  med_pay21 = median(reg_grosspaid21),
  entry21 = mean(FTE_pay21[EntryTERT21 == TRUE]),
  exper21 = mean(FTE_pay21[ExperTERT21 == TRUE]),
  avg_hours = mean(reg_hours),
  med_hours = median(reg_hours),
  q25_hours = quantile(reg_hours, .25),
  q75_hours = quantile(reg_hours, .75)
)

rm(nyc_group)

```

## Zeroes

*NOTE!!* There are zeroes for "Experienced" for titles with a very small n, because of the way I coded it (using a logical vector).
 
# Export

*NOTE* This code creates a csv. In the final versions I upload, I open the csv in Excel, convert to xlsx, format the data as a table, re-format text so currency variables and numbers are formatted correctly, move the file to NYC_Payroll/output, then upload to SharePoint.

```{r eval=FALSE}

write_csv(nyc_wages, "nyc_wages_byAgency_FY20adj21_09-29-21.csv")

```