---
title: "LMIS_Titles"
author: "Sam Neylon"
date: "August 4, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#### Using "here" to keep project directory as working directory
#### NOTE: Rmarkdown will try to make the file it is inside the working directory, but "here" should fix it.

library(here)
library(tidyverse)
library(data.table)

```

# FY20-21(adjusted) - Title Clusters

This file produces a summary statistics for titles in the NYC OpenData payroll information.

NYC OpenData: https://data.cityofnewyork.us/City-Government/Citywide-Payroll-Data-Fiscal-Year-/k397-673e

Notes:
  * This file uses the raw data for FY2020
  * This file then creates 2021 statistics by *1.7%*
    * This is in line with the adjustments made by the OEWS for state and local government workers: https://statistics.labor.ny.gov/lstechoes.shtm
  * In line with OEWS statistics, this file estimates two wage statistics:
    1. "Entry" - this is the average wage for the bottom 1/3 of wage earners in that title
    2. "Experienced" - this is the average wage for the top 2/3 of wage earners in that title.
  * Annualized wages
    * People who worked less than a year were dragging down averages
    * To compensate for this, we decided to annualize wages, but using information on how many annual hours a title typically worked
    * See "Annualizing Wages" section for more information.
    * *NOTE* statistics for the whole workforce were annualized, but for Parks, an exception was made for part-time workers.

## Title Clusters

To get a more accurate sense of the wage range of a title, we clustered similar titles together. For example, the wage range for a "Landscape Architect" doesn't just encompass the actual title "Landscape Architect", but also positions above and below it (i.e. Assistant Landscape Architect).

The goal of this is to make the data more comparable to the OEWS data, which gives you wage statistics by *occupation*, which may encompass several actual City titles.

Using DC 37 provided title descriptions, job ads, and civil service descriptions, we created lists of related titles.

In this file, we will calculate summary wage statistics based on these title clusters for Parks and Recreation-related titles. Note that these aren't just titles within the Parks dept., but are titles selected by LMIS and DC 37.

*Reference* the file NYC_Payroll/doc/DC37_JobTitlesDetailedContent_Parks.xlsx
  * The sheet "Title Clusters" is where Eileen noted her choices of titles for clusters.

# Import

## fread

```{r eval=TRUE}

# Import/Rename columns

nyc_data <- fread(here("data/ALL_Citywide_Payroll_Data__Fiscal_Year_.csv"), col.names = c("FY", "payroll_num", "agency", "last", "first", "ME", "start_date", "borough", "title", "leave_status", "base_salary", "pay_basis", "reg_hours", "reg_grosspaid", "OT_hrs", "OT_paid", "total_otherPay"))

# Rename columns - Filter for FY2020

nyc_data <- nyc_data %>% select(FY, agency, title, base_salary, pay_basis, reg_hours, reg_grosspaid) %>% 
  filter(FY == 2020)


```

# Clean

* Filter out zero reg_grosspaid and zero reg_hours
* "hrly": Calculate "hourly" (not official hourly rate) by dividing Gross Paid by Regular (annual) hours
* Create an adjusted 2021 wage by multiplying regular gross paid by 1.7%

* Some job titles have asterisk or question marks - after cursory look at the data, I don't believe they are substantially different, so I delete these symbols.

```{r eval=TRUE}

nyc_data <- nyc_data %>%
  filter(reg_grosspaid > 0 & reg_hours > 0) %>% 
  mutate(
  hrly = reg_grosspaid / reg_hours,
  reg_grosspaid21 = reg_grosspaid * 1.017,
  hrly21 = reg_grosspaid21 / reg_hours
)

nyc_data <- nyc_data %>% mutate(
  title = gsub("\\*", "", title),
  title = gsub("\\?", "", title)
)

```

# Annualizing Wages

There is one important issue with using "grosspaid" in this dataset. This is that if employees leave before a full year, their grosspaid will be very low, and thus will drag down the summary statistics that we are calculating.

We could simply take the grosspaid and divide it by the hours worked in a year (reg_grosspaid / reg_hours), which is how we calculate the "hrly" variable, and annualize it by multiplying by a set number of annual hours. 

*However*, some positions actually *are* seasonal, or work less than 2,080 hours a year (40 hrs a week * 52 weeks), so this would remove important information about the wages of certain titles.

We have settled on the following solutions:

  Seasonal: Through reading qualitative sources (union job descriptions, civil service listings) we may determine that a particular job title is seasonal. If so, we will treat it as seasonal in code which is tailored to a specific cluster of job titles (see "Parks_wages_titleClusters").

## Everything else: 

* Create new variable "FT_hours"
  * Calculate the 75th percentile of annual hours in a given Fiscal Year for a given Title.
  * If the 75th percentile is <= 1,820 hours, then set the FT_hours at 1,820 for this row (person-year).
  * If the 75th percentile is > 1,820 hours, then set FT_hours for this row to the 75th percentile for the title.
  
* Create new variable "FTE_pay"
  * If reg_hours >= FT_hours then set FTE_pay to reg_grosspaid
  * Else set FTE_pay to reg_grosspaid / reg_hours x FT_hours
  
## Parks Specific

Because Recreation careers often start with seasonal work, we have decided to *not* annualize the jobs going into "Recreation PT", or rather, we will use reg_hours and not FT_hours.


```{r eval=TRUE}

nyc_data <- nyc_data %>% group_by(title) %>% mutate(hrq75 = quantile(reg_hours, .75)) %>% ungroup()

nyc_data <- nyc_data %>% mutate(FT_hours = ifelse(hrq75 <= 1820, 1820, hrq75),
                                FTE_pay = ifelse(reg_hours >= FT_hours, reg_grosspaid, 
                                                 hrly * FT_hours),
                                FTE_pay21 = FTE_pay * 1.017)

```

# Parks Title Clusters Code

## Filter Data for only Parks and Rec brochure titles:

First I filter for just the titles we are using in the DC 37 Parks and Rec brochure.

I used the file NYC_Payroll/doc/Parks_Titlecluster_pre-process.xlsx to get the titles in a list with quotes around them, and comma delimited. This is so I can save them as vectors, which I can reference more easily in my code.

I will filter out CITY PARK WORKER titles who work less than 1,560 hours.

```{r eval=TRUE}

## All Parks titles

parks <- c("PARK SUPERVISOR", "PRINCIPAL PARK SUPERVISOR", "GARDENER", "CLIMBER & PRUNER", "ASSOCIATE PARK SERVICE WORKER", "CITY PARK WORKER", "RECREATION SUPERVISOR", "RECREATION SPECIALIST", "RECREATION DIRECTOR", "PLAYGROUND ASSOCIATE", "ASSISTANT LANDSCAPE ARCHITECT", "LANDSCAPE ARCHITECT", "LANDSCAPE ARCHITECT INTERN", "ADMINISTRATIVE HORTICULTURIST", "ADMINISTRATIVE PARKS & RECREATION MANAGER", "FORESTER", "ASSOCIATE URBAN PARK RANGER", "URBAN PARK RANGER")

## Title Clusters

Park_Supervisor <- c("PARK SUPERVISOR", "PRINCIPAL PARK SUPERVISOR")

Park_Service_Worker <- c("ASSOCIATE PARK SERVICE WORKER", "CITY PARK WORKER")

Landscape_Architect <- c("ASSISTANT LANDSCAPE ARCHITECT", "LANDSCAPE ARCHITECT", "LANDSCAPE ARCHITECT INTERN")

nyc_parks <- nyc_data %>% filter(title %in% parks) %>% filter(!(title %in% "CITY PARK WORKER" & reg_hours < 1560))


```

## Cluster ifelse

I will be using ifelse statements to group related titles [see methodology above]. For titles which aren't in clusters, they will be grouped by title.

I will slice the title "RECREATION SPECIALIST" into two titles, RECREATION PT and RECREATION FT, divided at 1560 annual hours (30hrs a week) - I came to this number using histograms to see the distribution of hours per employee in the title.

PLAYGROUND ASSOCIATE will be recoded to "RECREATION PT"

"ASSOCIATE URBAN PARK RANGER" will be recoded to "URBAN PARK RANGER"

"URBAN PARK RANGER" will be recoded to "PARKS ENFORCEMENT OFFICER"

```{r eval=TRUE}

nyc_parks <- nyc_parks %>% mutate(
  title_cluster = ifelse((title %in% Park_Supervisor), "PARK SUPERVISOR",
                          ifelse((title %in% Park_Service_Worker), "PARK SERVICE WORKER",
                                 ifelse((title %in% Landscape_Architect), "LANDSCAPE ARCHITECT",
                                        ifelse((title %in% "RECREATION SPECIALIST" & reg_hours >= 1560), "RECREATION FT",
                                               ifelse((title %in% "RECREATION SPECIALIST" & reg_hours < 1560), "RECREATION PT",
                                                      ifelse((title %in% "PLAYGROUND ASSOCIATE"), "RECREATION PT", 
                                                             ifelse((title %in% "ASSOCIATE URBAN PARK RANGER"), "URBAN PARK RANGER",
                                                                    ifelse((title %in% "URBAN PARK RANGER"), "PARKS ENFORCEMENT OFFICER",
                                                                           title))))))))
)

```

### Recreation Part-time recode

Since the following code will use the FTE (full time equivalent) annualized variables we have calculated, I will create new variables which replace the FTE_pay for "Recreation PT" with the original "reg_grosspaid" variable.

```{r eval=TRUE}

nyc_parks <- nyc_parks %>% mutate(pFTE_pay = ifelse(title_cluster %in% "RECREATION PT", reg_grosspaid, FTE_pay),
                                  pFTE_pay21 = ifelse(title_cluster %in% "RECREATION PT", reg_grosspaid21, FTE_pay21))

```


## Full Time coding

There are a few positions in which we keep rows with Part-Time workers, to get a sense of the actual annual pay that people can expect in some of these jobs with lots of seasonal and part-time positions.

However, for other full-time jobs, those working less than a full year seem to be people who left the job early, rather than a reflection of expected annual pay. For these positions, I will be using the "base_salary" variable. This variable is less useful than regular gross paid, because it reflects per hour pay for hourly workers, and also it reflects annualized pay for part-time workers who may be truly part time.

For observations which are not in the part-time careers we chose, who worked less than 1,560 hours, and who are paid on a per Annum basis, I will be replacing gross paid with base salary. 

# Summary Table

In line with OEWS statistics, this file estimates two wage statistics:
  1. "Entry" - this is the average wage for the bottom 1/3 of wage earners in that title
  2. "Experienced" - this is the average wage for the top 2/3 of wage earners in that title.
  
## First Tertile
  
  * First, calculate where the first third of wages within each title falls (the first tertile), and add this as a variable for each observation
  * Second, create two logical variables. Variable "Entry" is TRUE if a given worker's regular gross pay is less than or equal to the first tertile and FALSE if not. For "Exper" (experienced), TRUE if more than first tertile, FALSE if not.


```{r eval=TRUE}

nyc_group <- nyc_parks %>% group_by(title_cluster)

nyc_group <- nyc_group %>% mutate(tert = quantile(pFTE_pay, 1/3), tert21 = quantile(pFTE_pay21, 1/3)) %>% 
  mutate(EntryTERT = (pFTE_pay <= tert),
         ExperTERT = (pFTE_pay > tert),
         EntryTERT21 = (pFTE_pay21 <= tert21),
         ExperTERT21 = (pFTE_pay21 > tert21))

```

# Conditional Summaries

Now that I have a logical vector for Entry and Experienced, I can multiply this by the individual values (see code below) so my summaries are only for particular workers within a title.

NOTE: I have commented out two lines which gave me a count of how many Entry and Experienced in each position (I did this as a check on my work)

```{r eval=TRUE}

nyc_parks_wages <- nyc_group %>% summarise(
  tot_emp = n(),
  #entry_emp = sum(EntryTERT),
  #exper_emp = sum(ExperTERT),
  entry = mean(pFTE_pay[EntryTERT == TRUE]),
  exper = mean(pFTE_pay[ExperTERT == TRUE]),
  entry21 = mean(pFTE_pay21[EntryTERT21 == TRUE]),
  exper21 = mean(pFTE_pay21[ExperTERT21 == TRUE]),
  avg_hourly = mean(hrly),
  med_hourly = median(hrly),
  avg_hours = mean(reg_hours),
  med_hours = median(reg_hours),
  q25_hours = quantile(reg_hours, .25),
  q75_hours = quantile(reg_hours, .75),
  avg_pay = mean(reg_grosspaid),
  pay_q25 = quantile(reg_grosspaid, .25),
  med_pay = median(reg_grosspaid),
  pay_q75 = quantile(reg_grosspaid, .75),
  avg_pay21 = mean(reg_grosspaid21),
  pay21_q25 = quantile(reg_grosspaid21, .25),
  med_pay21 = median(reg_grosspaid21),
  pay21_q75 = quantile(reg_grosspaid21, .75)
)

rm(nyc_group)

```

 
# Export

*NOTE* This code creates a csv. In the final versions I upload, I open the csv in Excel, convert to xlsx, format the data as a table, re-format text so currency variables and numbers are formatted correctly, move the file to NYC_Payroll/output, then upload to SharePoint.

```{r eval=FALSE}

write_csv(nyc_parks_wages, "Parks_wages_titleClusters_FY20adj21_10-21-21.csv")

```





